---
title: "IGA-PLS"
author: "Zachary Kurtz"
output:
  pdf_document:
    highlight: tango
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

### Introduction
This Notebook is meant to contain the 16S_OTU data found within the respective manuscript. It contains only the code used to generate the figures found within the main text. This entire dataset is publicly available in QIITA under the ID [10527](https://qiita.ucsd.edu/study/description/10527). More details about the procedures used to generate the data can be found within the **Methods** section of the manuscript. The table found within the data folder has been processed to include consensus set of taxa.

#### 1a. Install and load the necessary libraries
```{r library, message = FALSE, warnings = FALSE}
library(stringr)
library(phyloseq)
# devtools::install_github('zdk123/compPLS')
library(compPLS)
library(spls)
```
#### 1b. Import the OTU and sample data
```{r import,  message = FALSE, warnings = FALSE}
pattern  <- "(.*)_cons\\.biom"
pattern2 <- paste('data', pattern, sep="/")
files    <- list.files('data', pattern, full.names=TRUE)
phylo_list <- lapply(files, import_biom)
names(phylo_list) <- str_match(files, pattern2)[,2]
```

#### prep data for PLS model ####
```{r prep, message=FALSE, warnings=FALSE}
patY <- sample_data(phylo_list$ppat)
Y <- cbind(as.numeric(patY$IgA),NA)
Y.split <- compPLS:::.Split.variation.two.level(Y,
       as.factor(patY$Treatment), as.factor(patY$Diet),
       as.factor(patY$sample_name))
X <- t(clr(phylo_list$ppat@otu_table@.Data, 2))
X.split <- compPLS:::.Split.variation.two.level(X,
       as.factor(patY$Treatment), as.factor(patY$Diet),
       as.factor(patY$sample_name))
Y <- scale(Y.split$Xw[,1,drop=FALSE], scale=FALSE, center=FALSE)
X <- scale(X.split$Xw,                scale=FALSE, center=TRUE)


```

#### run model, fit to transfer experiments ####
```{r fitpls, message=FALSE, warnings=FALSE}

source('scripts/trans_proc.R')

K <- 2
out.stars <- compPLS:::spls.stars(X, Y, rep.num=100, K=K,
             eta=seq(.799, .999, length.out=10), ncores=4)


p <- ncol(X)
threshs <- seq(0,1,.05)
n <- length(threshs)
out.spls_list <- vector('list', n)
Ypred_list    <- vector('list', n)
rsquared_in   <- vector('numeric', n)
rsquared_out  <- vector('numeric', n)
ntax <- vector('numeric', n)

for (i in 1:n) {
    th      <- threshs[i]
    ind     <- which(out.stars$merge[,1,1] >= th)
    ntax[i] <- length(ind)
    out.spls_list[[i]] <- spls::spls(X[,ind], Y, eta=0,
                          K=ifelse(K<ntax[i], K, ntax[i]), scale.x=TRUE, scale.y=FALSE)
    out.spls         <- out.spls_list[[i]]
    Ypred_list[[i]]  <- X[,ind] %*% out.spls$betahat[,1,drop=FALSE]
    rsquared_in[i] <- cor(Y, Ypred_list[[i]])^2
    Ytranspred <- Xtrans[,ind,drop=FALSE] %*% out.spls$betahat[,1,drop=FALSE]
    rsquared_out[i] <- cor(Ytrans[,1], Ytranspred[,1])^2
}

```
